---
title: "Capture recapture analyses of greens for Agnese Mancini"
output: html_notebook
---

This document describes data analysis of Agnese Mancini's data of green turtle capture recapture events of green  turtles using the augmented method (Kery and Schaub). To save computational time, I used MARK to select the best model (phiTSM, pt), and run just one model in this analysis. 


Initialize the workspace
```{r}
rm(list=ls())
library(jagsUI)
library(tidyverse)
library(lubridate)
library(reshape)
library(bayesplot)
library(ggridges)

source("Mancini_functions.R")

MCMC.params <- list(n.samples = 50000,
                    n.burnin = 30000,
                    n.thin = 5,
                    n.chains = 5)

```

Bring in the data file and get ready for CJS

```{r}

dat.1.Cm <- get.data.Cm("data/GTC_Cm Data_updated_2020-04-28_TE_v2.csv")
#dat.1 <- get.data("data/GTC_20190725_Tomo_v2.csv")

CJS.data <- dat2CJS(dat.1.Cm, save.file = FALSE)

CJS.data$data %>% rownames_to_column(var = "ID") -> data.CJS

# Number of years between occasions. 
cap.dates <- paste0(colnames(CJS.data$data), "-01")
delta.dates <- signif(as.numeric(as.Date(cap.dates[2:length(cap.dates)]) -
                                   as.Date(cap.dates[1:(length(cap.dates)-1)]))/365, 1)

# capture history
CH <- as.matrix(CJS.data$data)

nInd <- nrow(CJS.data$data)

ns <- colSums(CJS.data$data)

# find the first capture date
get.first <- function(x) min(which(x != 0))
f <- apply(CH, 1, get.first)

jags.data <- list(y = CH, 
                  f = f, 
                  nind = nInd, 
                  n.occasions = dim(CH)[2],
                  n = ns, 
                  T = ncol(CJS.data$data),
                  dt = c(0, delta.dates))


# initial values - when removing z, no input
# initsFunction <- function(ch){
#   mean.phi <- rbeta(1,2,2)
#   mean.p <- rbeta(1,2,2)
#   z <- known.state.cjs(ch)
#   b <- rnorm(1, 0, 50)
#   A <- list(mean.phi = mean.phi, mean.p = mean.p, z = z, beta = b)
#   return(A)
# }
```


Then run JAGS:

Time-since-marking - looking at the transients effect on survival 
```{r}

if (!file.exists(paste0("RData/CJS_PhiTSM_pt_N_logistic_Cm_JAGS.rds"))){
  MCMC.params$model.file <- "models/Model_CJS_PhiTSM_pt_N_logistic.txt"
  
  first.cap <- jags.data$f
  m <- matrix(data = 2, nrow = nrow(CH), ncol = ncol(CH))
  for (k in 1:nrow(m)){
    m[k, first.cap[k]] <- 1
  }
  
  jags.data$m <- m
  
  ## parameters to monitor - when this is changed, make sure to change
  ## summary statistics indices at the end of this script. 
  parameters <- c("gamma", "p", "N", "prop.trans", 
                  "mu_N", "sigma_N", "r", "K",
                  "deviance", "loglik")
  
  jm.PhiTSM.pt.N.logistic <- jags(data = jags.data,
                                  #inits = inits,
                                  parameters.to.save= parameters,
                                  model.file = MCMC.params$model.file,
                                  n.chains = MCMC.params$n.chains,
                                  n.burnin = MCMC.params$n.burnin,
                                  n.thin = MCMC.params$n.thin,
                                  n.iter = MCMC.params$n.samples,
                                  DIC = T, 
                                  parallel=T)
  
  out.list <- list(jags.out = jm.PhiTSM.pt.N.logistic,
                   jags.data = jags.data)
  
  saveRDS(out.list, 
          file = paste0("RData/CJS_PhiTSM_pt_N_logistic_Cm_JAGS.rds"))
} else {
  phiTSM.pt.N.logistic <- readRDS(file = "RData/CJS_PhiTSM_pt_N_logistic_Cm_JAGS.rds")

}
```


START HERE AFTER RUNNING JAGS - 2020-08-28

```{r}
mcmc_dens(phiAgeClass.pt$jags.out$samples, c("gamma[1]", "gamma[2]"))
```


```{r}
gamma.stats <- phiAgeClass.pt$jags.out$summary[grep("gamma", 
                                               dimnames(phiAgeClass.pt$jags.out$summary)[[1]]),]
gamma.stats
```


```{r}
p.stats <- phiAgeClass.pt$jags.out$summary[grep("p[", 
                                           dimnames(phiAgeClass.pt$jags.out$summary)[[1]],
                                           fixed = TRUE),]
p.stats
```


Survival rates of residents were about 0.6/yr.  Kinda low... 

Using the H-T estimator, we also computed the abundance for each sampling occasion:

```{r}
N.stats <- phiAgeClass.pt$jags.out$summary[grep("N[", 
                                           dimnames(phiAgeClass.pt$jags.out$summary)[[1]],
                                           fixed = TRUE),]

N.stats.df <- data.frame(N.stats)

N.stats.df$season <- colnames(CH.2)

N.stats.df
```


```{r}
ggplot(data = N.stats.df) +
  geom_point(aes(x = season, y = X50.)) + 
  geom_errorbar(aes(x = season, ymin = X2.5., ymax = X97.5.)) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 0.5, vjust = 0.5)) + 
  ylab("Abundance") + xlab("")

```

