---
title: "Capture recapture analyses for Agnese Mancini"
output: html_notebook
---

This document describes data analysis of Agnese Mancini's data of green turtle capture recapture events. 


Initialize the workspace
```{r}
rm(list=ls())
library(jagsUI)
library(tidyverse)
library(lubridate)
library(reshape)
library(bayesplot)
library(ggridges)
library(RMark)
library(R2ucare)

source("Mancini_functions.R")

do_analysis <- function(dp, ddl)
{
  # create formulas for Phi
  # tsm is time-since-marking; check for transient effects
  Phi.dot <-  list(formula = ~ 1)  
  #Phi.weight <- list(formula= ~ min_weight)   # many missing data 
  #Phi.t <- list(formula = ~ time)             # we never have this model worked for turtles... 
  #Phi.season <- list(formula = ~ sum_win)      # this also is unlikely... 
  #Phi.transience <- list(formula = ~ Transient)
  Phi.tsm <- list(formula = ~ tsm)
  
  #create formulas for p
  p.dot <- list(formula = ~ 1)
  p.t <- list(formula = ~ time)
  p.tsm <- list(formula = ~ tsm)
  #p.transience <- list(formula = ~ Transient)
  #p.tsm.transience <- list(formula = ~ tsm + Transient)
  #p.t.transience <- list(formula = ~ time + Transient)
  p.effort <- list(formula = ~ effort)
  p.season <- list(formula = ~ sum_win)
  p.tsm.season <- list(formula = ~ tsm + sum_win)
  p.tsm.effort <- list(formula = ~ tsm + effort)
  
  # create all combinations 
  cml <- create.model.list("CJS")
  
  # run all all models and return as a list with class marklist
  results <- mark.wrapper(cml,
                          data=dp,
                          ddl=ddl,
                          output=FALSE,
                          silent=TRUE)
  return(results)
}

```

Bring in the data file and get ready for CJS

```{r warning=F}
#dat.1 <- get.data("data/GTC_20190725_Tomo_v2.csv")

dat.1.Cm <- get.data.Cm("data/GTC_Cm Data_updated_2020-04-28_TE_v2.csv")

# there are In in the species code... 
#dat.1 %>% filter(species == "Cm") -> dat.1.Cm

CJS.data <- dat2CJS(dat.1.Cm, save.file = FALSE)

CJS.data$data %>% rownames_to_column(var = "ID") -> CH.1 #data.CJS

# using R2ucare::group_data to combine CHs
CH.2 <- R2ucare::group_data(CJS.data$data, rep(1, nrow(CJS.data$data)))

# remove weight because not all turtles were weighed...
# dat.1.Cm %>% select(ID, weight_kg) %>% group_by(ID) %>%
#   summarise(min_weight = min(weight_kg, na.rm = T)) %>%
#   filter(!is.infinite(min_weight)) -> data.weight
# 
# data.weight  %>% left_join(data.CJS, by = "ID") %>%
#   select(-c("ID", "min_weight")) -> CH.1
# 
# data.weight  %>% left_join(data.CJS, by = "ID") %>%
#   select(min_weight) -> cov.weight

# here I define transients to be those that were caught just once
# but that eliminates the possibility of dying after the first
# capture - how do I deal with this?
# n.cap <- rowSums(CJS.data$data)
# transient.vec <- rep(1, times = length(n.cap))
# transient.vec[n.cap > 1] <- 2

# need to count how many capture events occurred per season
dat.1.Cm %>% select(season, "DATE") %>% 
  group_by(season) %>% #-> tmp3
  summarise(effort = n_distinct(DATE)) -> effort.season

# capture history
tmp <- apply(as.matrix(CH.2[, 1:(ncol(CH.2)-1)]), 
             MARGIN = 2,
             FUN = paste0)

CH <- unite(data.frame(tmp),
           col = "ch",
           sep = "")

CH$freq <- CH.2$effY

# tmp <- apply(as.matrix(CJS.data$data), 
#              MARGIN = 2, 
#              FUN = paste0)

# CH <- unite(data.frame(tmp), 
#             col = "ch", 
#             sep = "")

#CH$min_weight <- cov.weight
#CH$Transient <- transient.vec


# capture dates and difference in years
cap.dates <- paste0(colnames(CJS.data$data), "-01")
delta.dates <- signif(as.numeric(as.Date(cap.dates[2:length(cap.dates)]) -
                                   as.Date(cap.dates[1:(length(cap.dates)-1)]))/365, 1)

# a possibility of survival changing between summer/winter?
tmp <- strsplit(colnames(CJS.data$data), split = "-")
tmp2 <- lapply(tmp, FUN = function(x) x[2])
tmp3 <- unlist(tmp2)

# dp <- process.data(CH, 
#                    model = "CJS", 
#                    time.intervals = delta.dates,
#                    groups = "Transient",
#                    begin.time = 2001)

dp <- process.data(CH, 
                   model = "CJS", 
                   time.intervals = delta.dates,
                   begin.time = 2001)

ddl <- make.design.data(dp)

# p is indexed from 2001.4
# Phi is indexed from 2001

# effort affects the capture probability (p)
effort.season <- effort.season[1:length(levels(ddl$p$time)),]
effort.season$time <- as.factor(2001 + cumsum(delta.dates))

ddl$p <- merge_design.covariates(ddl$p, df = effort.season)

# summer/winter affects capture probability (maybe)
sum_win <- data.frame(sum_win = unlist(ifelse(tmp3 == "08", "summer", "winter")) [1:length(levels(ddl$p$time))],
                      time = as.factor(2001 + cumsum(delta.dates)))

ddl$p <- merge_design.covariates(ddl$p, df = sum_win)

# summer/winter affects survival (maybe)
sum_win <- data.frame(sum_win = unlist(ifelse(tmp3 == "08", "summer", "winter")) [1:length(levels(ddl$p$time))],
                      time = as.factor(2001 + cumsum(c(0, delta.dates[1:(length(delta.dates)-1)]))))

ddl$Phi <- merge_design.covariates(ddl$Phi, df = sum_win)

# add time-since-marking (TSM models) - not sure if I'm doing this right... 
# When age = 0, it is the first time turtles were caught group(1)
# all others are group(2), which are recaptures. 
ddl$Phi$tsm <- 1
ddl$Phi$tsm[ddl$Phi$age == 0] <- 2

ddl$p$tsm <- 1
ddl$p$tsm[ddl$p$age == 0] <- 2

Cm.results <- do_analysis(dp = dp, ddl = ddl)

Cm.CJS.input <- list(raw.data = dat.1.Cm,
                     CJS.data = CJS.data,
                     CH.1 = CH.1,
                     CH.R2ucare = CH.2,
                     effort = effort.season,
                     CH = CH,
                     delta.dates = delta.dates,
                     dp = dp,
                     ddl = ddl,
                     run.date = Sys.Date())

saveRDS(Cm.results, file = "RData/CJS_Cm_RMark.rds")
saveRDS(Cm.CJS.input, file = "RData/CJS_Cm_RMark_input.rds")

```

Export to Mark to see Release GOF or median c-hat.
```{r}
export.MARK(dp, 
            project.name = "Mancini_Cm", 
            model = Cm.results$Phi.tsm.p.t, 
            replace = T, 
            chat = 5.6)
```

Median c-hat test returned semi-acceptable results of estimated c-hat = 1.xxx, with 95%CI = ().  

Release output is similar to what I find below from R2ucare.

Compare results using AICc:

```{r}
model.table(Cm.results)
```

According to AICc, the best one is Phi(tsm)p(time). But c-hat > 4 (4.75)!  This is better than before but, obviously, this isn't a great model... What to do? 

Look at the estimates:
```{r}
Cm.results$Phi.tsm.p.t$results$real
```

